Liveness-Based Garbage Collection for Lazy Languages

### 2015-10-06 paper/paper.tex
Moved GC to start of function calls. Need to fix errors

The number  of garbage  collections is same  since we  approximate the
liveness of  a closure  to the  liveness at  its creation  point.  The
liveness  we get  is usually  grossly over  approximated.  Using  this
liveness may not  give too much of a benefit  compared to reachability
as nearly everything would be live at its point of creation.

@@

Memory   usage   graphs   for   select   benchmarks   in   are   shown
Figure~\ref{fig:memory-usage}, Column  (a).  As the number  of garbage
collections tend to  be very large, for each benchmark  we also show a
window which  is representative  of the  behavior for  that particular
benchmark (Column (b)).  In all the programs we can see that the curve
corresponding to  LGC (red  line) regularly dips  below the  RGC curve
(blue line), the dip is specially visible for the benchmarks where LGC
outperforms  RGC substantially.   The  graphs also  include curve  for
reachable cells (black,  obtained approximately by forcing  RGC to run
at a very high frequency) and the live cells for particular run (lightblue,
obtained by post processing the heap access at the end of a program).

@@

%% One issue we  highlighted was that while fewer  nodes were marked (and
%% more garbage collected), sometimes  cons cells could be visited and
%% traversed  multiple times  with different  sets of  liveness  paths to
%% explore.  Further work includes  improvements to the classical copying
%% collector to reduce the cost of this.
We extended the liveness based garbage collection to lazy
  languages and shown its benefit  for garbage collection in practical
  programs. We  defined context  sensitive liveness analysis  that uses
  context independent summaries of functions obtained using a symbolic
  demand. We show  that obtaining precise solution  to these equations
  is undecidable in  our formulation, and hence  we safely approximate
  the result using DFA. These DFA are consulted by the garbage collector
  to improve  collection.

  Liveness   of   closure   presented  further   challenges   to   our
  implementation.   We compared  different approximate  strategies for
  handling   liveness   of   closures,   and  found   that   a   mixed
  strategy---reachability  based  collection   of  closure  arguments,
  liveness  based  collection  of   everything  else---works  best  in
  practice as it  avoids runtime and space  overheads.  With mixed-mode
  garbage collection scheme,  we were able to collect  more cells than
  reachability based collectors at a reasonable overhead.

>>>>

The number  of garbage  collections is the same  since we  approximate the
liveness of  a closure  to the  liveness at  its creation  point.  The
liveness  we get  is usually  grossly over  approximated.  Using  this
liveness may not  give too much of a benefit  compared to reachability
as nearly everything would be live at its point of creation.

@@

Memory   usage   graphs   for   select   benchmarks    are   shown
Figure~\ref{fig:memory-usage}, Column  (a).  As the number  of garbage
collections tend to  be very large, for each benchmark  we also show a
window which  is representative  of the  behavior for  that particular
benchmark (Column (b)).  In all the programs we can see that the curve
corresponding to  LGC (red  line) regularly dips  below the  RGC curve
(blue line), the dip is specially visible for the benchmarks where LGC
outperforms  RGC substantially.   The  graphs also  include curve  for
reachable cells (black,  obtained approximately by forcing  RGC to run
at a very high frequency) and the live cells for particular run (lightblue,
obtained by post processing the heap access at the end of a program).

@@

%% One issue we  highlighted was that while fewer  nodes were marked (and
%% more garbage collected), sometimes  cons cells could be visited and
%% traversed  multiple times  with different  sets of  liveness  paths to
%% explore.  Further work includes  improvements to the classical copying
%% collector to reduce the cost of this.
We extended the liveness based garbage collection to lazy
  languages and shown its benefit  for garbage collection in practical
  programs. We  defined a context  sensitive liveness analysis  that uses
  context independent summaries of functions obtained using a symbolic
  demand. We showed  that obtaining precise solution  to these equations
  is undecidable in  our formulation, and hence  we safely approximate
  the result using DFA. These DFA are consulted by the garbage collector
  to improve  collection.

  Liveness   of   closures   presented  further   challenges   to   our
  implementation.   We compared  different approximate  strategies for
  handling   liveness   of   closures,   and  found   that   a   mixed
  strategy---reachability  based  collection   of  closure  arguments,
  liveness  based  collection  of   everything  else---works  best  in
  practice as it  avoids runtime and space  overheads.  With mixed-mode
  garbage collection scheme,  we were able to collect  more cells than
  reachability based collectors at a reasonable overhead.
